---
title: "Reusable PubMed Search Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reusable PubMed Search Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

This vignette shows how to use `pubmedsearchworkflow` to keep PubMed systematic
searches reusable across projects. The package separates study-specific search
strings from the R code that reads, chunks, runs, combines, and exports PubMed
records.

The bundled quaternary ammonium compound (QAC) search is an example study. Use
its file layout as a template for new searches rather than editing the package
functions.

## Install and Load

Install dependencies and this package from a fresh R session opened at the
repository root. A fresh session matters on Windows because packages such as
`cli`, `fs`, and `rlang` can be locked if they are already loaded by RStudio or
another package.

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_deps(dependencies = TRUE, upgrade = "never")
system2(file.path(R.home("bin"), "R"), c("CMD", "INSTALL", "."))
```

If installation fails with a message such as `cannot remove prior installation`
or `namespace 'fs' ... is being loaded`, restart R and run the same commands
before loading `devtools`, `pkgdown`, or this package.

Load the package:

```{r}
library(pubmedsearchworkflow)
```

## Locate Example Search Files

When working from the source checkout, paths such as `inst/examples/...` are
available directly. After installation, use `system.file()` to locate bundled
examples inside the installed package.

```{r}
qac_include <- system.file(
  "examples/qac/search_terms/qac_chemical_string_841.txt",
  package = "pubmedsearchworkflow"
)

qac_exclude <- system.file(
  "examples/qac/search_terms/qac_exclusion_string.txt",
  package = "pubmedsearchworkflow"
)

smoke_query <- system.file(
  "examples/smoke/pubmed_pmid_query.txt",
  package = "pubmedsearchworkflow"
)

file.exists(c(qac_include, qac_exclude, smoke_query))
```

## Read and Normalize Queries

Study teams should store PubMed search strings in plain-text files. This makes
review, version control, and reuse easier than embedding long Boolean strings in
notebooks.

```{r}
include_query <- read_pubmed_query(smoke_query)
include_query
```

`read_pubmed_query()` collapses line breaks and repeated whitespace so query
files can be formatted for readability.

```{r}
normalize_pubmed_query(c('"alpha"[Title]', " OR ", ' "beta"[Title]'))
```

## Add Exclusions

Use `build_pubmed_query()` to combine an include query with optional exclusions.
The helper wraps both sides in parentheses to preserve intent.

```{r}
build_pubmed_query(
  include_query = '"alpha"[Title]',
  exclude_query = '"review"[Publication Type]'
)
```

## Split Large Top-Level OR Queries

Large chemical-name searches often need to be split into smaller PubMed requests.
`split_pubmed_or_query()` splits only top-level `OR` operators. It preserves
quoted strings and nested parenthetical logic.

```{r}
example_query <- '("a"[Title] AND ("b"[Title] OR "c"[Title])) OR "d OR e"[Title] OR "f"[Title]'
split_pubmed_or_query(example_query)
```

For the QAC example, the large include file can be split into chunks before
submitting to PubMed.

```{r}
qac_chunks <- split_pubmed_or_query(read_pubmed_query(qac_include))
length(qac_chunks)
head(qac_chunks, 2)
```

## Run a Live PubMed Search

Live PubMed calls are intentionally not evaluated in this vignette so package
checks remain deterministic. Run this smoke test manually when you want to
verify local network access and `easyPubMed` behavior.

```{r, eval = FALSE}
results <- run_pubmed_search(
  include_query = read_pubmed_query(smoke_query),
  chunk = FALSE
)

records <- combine_pubmed_data(results)
records[, c("pmid", "title", "year", "jabbrv")]
```

For a broad query, start with a small chunk limit before launching a full search:

```{r, eval = FALSE}
results <- run_pubmed_search(
  include_query = read_pubmed_query(qac_include),
  exclude_query = read_pubmed_query(qac_exclude),
  chunk = TRUE,
  limit = 2,
  sleep = 1
)

records <- combine_pubmed_data(results)
```

Remove `limit = 2` only after the smoke run succeeds.

## Export Results

Use `write_pubmed_csv()` to combine records, deduplicate by PMID, create parent
directories, and write a CSV.

```{r, eval = FALSE}
write_pubmed_csv(results, "outputs/qac_search_smoke.csv")
```

## Command-Line Runner

Collaborators who do not need to write R code can use the bundled runner from
the repository root:

```{r, eval = FALSE}
system2(
  "Rscript",
  c(
    "exec/run_pubmed_search.R",
    "--include", "inst/examples/smoke/pubmed_pmid_query.txt",
    "--output", "outputs/smoke_pubmed_pmid.csv",
    "--no-chunk"
  )
)
```

The runner uses the same package functions as the R API. Use `--limit` for
smoke tests, `--sleep` to pause between PubMed requests, and `--quiet` to
suppress package-level progress messages.

## Recommended Project Pattern

For a new systematic search:

1. Copy `inst/examples/qac/` to a new project-specific folder.
2. Replace the include and exclusion text files with the new search strategy.
3. Run a one-PMID or low-limit smoke test.
4. Commit search-term files and workflow notes.
5. Keep generated CSV files in `outputs/` unless they are final curated
   deliverables.
